Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-03-04 paper/main.tex
section 4 completed

\usepackage{multirow}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{makecell}
\usepackage[table]{xcolor}
%\usepackage{arydshln}
%\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}

@@

\section{Reduction algorithm}\label{sec:algo}
Jak se nastavuji inicialni parametry a jak postupne klesaji.
Jak prechazim od nizsich levelu k vyssim.
Vyuzivame incremental QBF solving?

\todo{Nasleduje stary text, ale pozor, obrazek je novy!}
Now we describe our simplification algorithm taking as input an arbitrary TELA $\mathcal{A}$ with an acceptance condition $\psi$, required level of simplification $L$ and timeout period $T$.
\begin{enumerate}
    \item Convert acceptance condition $ \psi$ to DNF.
    \item Apply the algorithm for simplification Level 1 described in Section \ref{sec:standard} on $\mathcal{A}$.
    \item If $L = 1$ return automaton $\mathcal{A}$.
     \item The initial value of $C$ is the number of clauses in $\psi$. Similarly, the initial value of $K$ is the number of acceptance sets of $\psi$ decreased by 1.
    \item Encode $\mathcal{A}$ into QBF formula $\varphi$ according to the required level of simplification $L$.
    \item Run a QBF-solver on $\varphi$ for $T$ seconds.
    \begin{enumerate}
        \item If the QBF-solver evaluates $\varphi$ as UNSAT, return automaton $\mathcal{A}'$.
        \item If the timeout $T$ expires before the QBF-solver evaluates $\varphi$, return automaton $\mathcal{A}'$.
        \item If the QBF-solver evaluates $\varphi$ as SAT. Decode the assignment into next TELA $\mathcal{A}'$.
        \begin{enumerate}
            \item If $K=1$, try to reduce $\varphi$ to $\true$ or $\false$, the simplest form of acceptance condition and return $\mathcal{A}$.
            \item Decrease $K$ by 1 and go to (b).

    \end{enumerate}

    \end{enumerate}
\end{enumerate}

\todo{kde je optimalni misto, kam napsat:}
For simplicity, we do not display letters on the edges of the automata in Figures. We can afford to omit them because they are not important for our simplification techniques.

\begin{itemize}
\item obecny popis konstrukce formule
\item (level 3)
\item level 4
\end{itemize}

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[smallautomaton]
    \begin{scope}[initial angle=90,xshift=-3.5cm]
      \node (l) at (1,1.3) {input:};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={1}{1}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={2}{2}} pic[pos=.67] {acc={3}{3}} pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {acc={3}{3}} pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {\small$(\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{3}{3})$};
    \end{scope}
    \begin{scope}[initial angle=90,xshift=3.5cm]
      \node (l) at (1,1.3) {level 1:};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={2}{2}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={0}{0}} pic[pos=.67] {acc={2}{2}} pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {acc={1}{1}} pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {\small$(\Fin\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Fin\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})$};
    \end{scope}
    \begin{scope}[initial angle=90,xshift=-3.5cm,yshift=-4cm]
      \node (l) at (1,1.3) {level 2:};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={0}{0}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic {acc={0}{0}} pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {acc={1}{1}} pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {\small$(\Fin\tacc{0}{0}\wedge\Inf\tacc{1}{1})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1})$};
    \end{scope}
    \begin{scope}[initial angle=90,xshift=3.5cm,yshift=-4cm]
      \node (l) at (1,1.3) {level 3:};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={0}{0}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {\small$\Fin\tacc{0}{0}$};
    \end{scope}
  \end{tikzpicture}
  \smallskip

  \caption{An example illustrating the results of the three basic
    reduction levels: an input automaton and the automata
    obtained by reduction of level 1, level 2, and level 3.
    \todo{Pozor, prvni klauzule v levelu 2 je trash!}}
  \label{fig:example}
\end{figure}

>>>>

\usepackage{multirow}
\usepackage{booktabs}
\usepackage{xspace}
\usepackage{makecell}
\usepackage[table]{xcolor}
%\usepackage{arydshln}
%\newcommand\VRule[1][\arrayrulewidth]{\vrule width #1}
\usepackage[noend]{algorithm2e}

@@

\section{Reduction algorithm}\label{sec:algo}

This section explains how we use the QBF formulas constructed in the
previous section to reduce the number of acceptance marks in TELA
automata. First we describe a \emph{single-level} reduction, which
uses only a single kind of QBF formulas. More precisely, we talk about
\emph{level~1} reduction if $\Phi^\mathit{opt}_{1,C,K}$ is
used. Analogously, we talk about \emph{level~2} and \emph{level~3}
reduction when $\Phi^\mathit{opt}_{2,C,K}$ or
$\Phi^\mathit{opt}_{3,C,K}$ is used, respectively.

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetKwFunction{FextendBackward}{extendBackward}
  \SetKwProg{PI}{Procedure}{}{}
  \PI{$\mathit{SingleLevelReduction}(\mA,j,\mathit{reduceC})$}{
    \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, level $j\in\{1,2,3\}$, $\mathit{reduceC}\in\{\true,\false\}$}
    \KwOut{an equivalent TELA with the same structure as $\mA$ and with at most as many acceptance marks as in $\mA$} %with the same or lower number of acceptance~marks}
    \;
    \If{$L(\mA)=\emptyset$}{
%      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'=\{(p,a,\emptyset,q)\mid(p,a,M',q)\in\delta\}$
      \KwRet $(Q,\emptyset,\Sigma,\delta',q_I,\false)$ where $\delta'$ is $\delta$ with all acceptance marks removed
    }
    $C_\mA\gets$ the number of clauses in the acceptance formula of $\mA$ transformed to DNF\;
    $K_\mA\gets$ the number of acceptance marks in $\mA$\;
    $C\gets C_\mA$\;
    $K\gets K_\mA$\;
    $\mathit{stop}\gets\false$\;
    \While{$\neg\mathit{stop}~\wedge~K>0$}{
      $\Phi\gets\Phi^\mathit{opt}_{j,C,K{-}1}$ for $\mA$\;
      \leIf{$\mathit{satisfiable}(\Phi)$}{$K\gets K{-}1$}{$\mathit{stop}\gets\true$}
    }
    \If{$\mathit{reduceC}$}{
      $\mathit{stop}\gets\false$\;
      \While{$\neg\mathit{stop}~\wedge~C>1$}{
        $\Phi\gets\Phi^\mathit{opt}_{j,C-1,K}$ for $\mA$\;
        \leIf{$\mathit{satisfiable}(\Phi)$}{$C\gets C{-}1$}{$\mathit{stop}\gets\true$}
      }
    }
    \eIf{$K=K_\mA~\wedge~C=C_\mA$}{
      \KwRet $\mA$
    }{
      compute $\rem$ and $\psi$ from a model of $\Phi^\mathit{opt}_{j,C,K}$\;
      \KwRet $(Q,N_K,\Sigma,q_I,\rem(\delta),\psi)$
    }
  }\;
  \caption{The single-level reduction procedure}
  \label{alg:single}
\end{algorithm}

The reduction procedure is given in Algorithm~\ref{alg:single}. Besides
the reduction of acceptance marks, the algorithm also reduces the
number of clauses in acceptance formula if the last argument
$\mathit{reduceC}$ is set to $\true$. In the first step, the procedure
checks whether the input automaton represents the empty language and
if this is the case, it returns the same automaton with acceptance
conditon $\false$ and with all acceptance marks removed. Otherwise it
gradually decreases the number of marks until 0 is reached or the QBF
solver behind the function $\mathit{satisfiable}(\Phi)$ fails to
reduce the number of marks, i.e., it claims unsatisfiability of $\Phi$
or it runs out of resources. In the next step, if $\mathit{reduceC}$
is set to $\true$ then the procedure gradually reduces the number of
clauses. It stops if no equivalent automaton with less clauses is
found or the number of clauses is reduced to one. Note that the
acceptance condition with 0 clauses is $\false$ and this case is
treated at the beginning of the procedure. Finally, if the procedure
succeeds to reduce the number of marks or clauses, if constructs the
modified automaton. Otherwise, it returns the original automaton.
\todo{do sekce o implementaci napsat, ze namisto opakovaneho budovani
  $\Phi_{j,C,K}$ puzivame inkrementalni pristup.}

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[smallautomaton]
    \begin{scope}[initial angle=90,xshift=-3.5cm]
      \node (l) at (1,1.3) {automaton $\mA$};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={1}{1}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={2}{2}} pic[pos=.67] {acc={3}{3}} pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {acc={3}{3}} pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {$(\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2})\vee(\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{3}{3})$};
    \end{scope}
    \begin{scope}[initial angle=90,xshift=3.5cm]
      \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,1,\true)$}; %{level 1:};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={2}{2}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic[pos=.33] {acc={0}{0}} pic[pos=.67] {acc={2}{2}} pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {acc={1}{1}} pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {$\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}\wedge\Inf\tacc{2}{2}$};
    \end{scope}
    \begin{scope}[initial angle=90,xshift=-3.5cm,yshift=-4cm]
      \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,2,\true)$}; %{level 2:};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={0}{0}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic {acc={0}{0}} pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {acc={1}{1}} pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {$\Inf\tacc{0}{0}\wedge\Fin\tacc{1}{1}$};
    \end{scope}
    \begin{scope}[initial angle=90,xshift=3.5cm,yshift=-4cm]
      \node (l) at (1,1.3) {$\mathit{SingleLevelReduction}(\mA,3,\true)$}; %{level 3:};
      \node[state,initial] (p) at (0,0) {};
      \node[state] (q) at (2,0) {};
      \path[->,auto]
      (p) edge[out=-150,in=150,looseness=10] pic {acc={0}{0}} pic {l=$a$} (p)
      (p) edge[out=30,in=150] pic {acc={0}{0}} pic {l=$b$} (q)
      (q) edge[out=30,in=-30,looseness=10] pic {l=$c$} (q)
      (q) edge[out=-150,in=-30] pic {l=$d$} (p);

      \node[acclabel] at (1,-1.4)
      {$\Fin\tacc{0}{0}$};
    \end{scope}
  \end{tikzpicture}
  \smallskip

  \caption{An example illustrating the results of the three basic
    reduction levels: an input automaton and the automata
    obtained by reduction of level 1, level 2, and level 3.}
  \label{fig:example}
\end{figure}

Figure~\ref{fig:example} shows a very simple automaton $\mA$ and the
three automata produced by calls of
$\mathit{SingleLevelReduction}(\mA,j,\true)$ for $j\in\{1,2,3\}$. The
figure clearly illustrates that the higher level of reduction we use,
the more acceptance marks can be reduced. On the other side, lower
levels are typically faster. The best results can be often achieved by
combining reductions of all levels. We call this aproach
\emph{multi-level} reduction. It is a straightforward sequential
application of the three levels, see Algorithm~\ref{alg:multi}.

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetKwFunction{FextendBackward}{extendBackward}
  \SetKwProg{PI}{Procedure}{}{}
  \PI{$\mathit{MultiLevelReduction}(\mA,\mathit{reduceC})$}{
    \KwIn{TELA $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$ and $\mathit{reduceC}\in\{\true,\false\}$}
    \KwOut{an equivalent TELA with the same structure as $\mA$ and with at most as many acceptance marks as in $\mA$} %with the same or lower number of acceptance~marks}
    \;
    $\mA\gets\mathit{SingleLevelReduction}(\mA,1,\mathit{reduceC})$\;
    $\mA\gets\mathit{SingleLevelReduction}(\mA,2,\mathit{reduceC})$\;
    $\mA\gets\mathit{SingleLevelReduction}(\mA,3,\mathit{reduceC})$\;
    \KwRet $\mA$
  }\;
  \caption{The multi-level reduction procedure}
  \label{alg:multi}
\end{algorithm}
